## MTE Project
Purpose: Metabolic rate and mass in ectotherms after 3* increase in temperature


#### CALCULATING EXPONENTS AND ACTIVATION ENERGY FOR METABOLIC EQUATION
This section uses metabolic rate, temperature, and size data to calculated fitted exponent and activation energy for the taxonomic classes for which we could find data. 

```{r, echo=FALSE}
# Classes: Branchiopoda, Malacostraca, Maxillopoda
# Data source: Makrieva
MR_data = read.csv("Class_metabolicrates_Makrievadata.csv", stringsAsFactors = FALSE)
Class_list=unique(MR_data$Class)
InvK=1/(273.15+MR_data$TC) #getting all the units correct
Mkg=MR_data$Mg/1000
W=MR_data$qWkg*Mkg
MR_data = cbind(MR_data, InvK, W)

#obtaining class-specific exponents for metabolic rate scaling for Branchiopoda Malacostraca and Maxillopoda
class_slopes=c()
for (current_class in Class_list){
  Class_data=subset(MR_data, MR_data$Class == current_class)
  basic=lm(log(Class_data$W) ~ log(Class_data$Mg) + Class_data$InvK)
  Mslope=basic$coefficients[2]
  Tslope=basic$coefficients[3]
  Ea=Tslope*0.00008617
  class_slopes=rbind(class_slopes, c(current_class, Mslope, Tslope, Ea))
}

# Classes: Amphibia
# Data source: White
MR_Amphi = read.csv("Whiteetal_Amphibiandata.csv", stringsAsFactors = FALSE)
A_InvK=1/(273.15+MR_Amphi$TC)
A_Mkg=MR_Amphi$Mg/1000
MR_Amphi = cbind(MR_Amphi, A_InvK, A_Mkg)
A_basic=lm(log(MR_Amphi$Watts) ~ log(MR_Amphi$A_Mkg) + MR_Amphi$A_InvK)
Mslope=A_basic$coefficients[2]
Tslope=A_basic$coefficients[3]
Ea=Tslope*0.00008617
class_slopes=rbind(class_slopes, c("Amphibia", Mslope, Tslope, Ea))
class_slopes=as.data.frame(class_slopes, stringsAsFactors = FALSE)

# Classes: Actinoperygii
# Data source: Gillooly
#obtaining super-class-specific exponents for metabolic rate scaling in fish
MR_fish = read.csv("gillooly_fish.csv", stringsAsFactors = FALSE)
f_Mkg=MR_fish$Mg/1000
MR_fish = cbind(MR_fish, f_Mkg)
F_basic=lm(log(MR_fish$W) ~ log(MR_fish$f_Mkg) + MR_fish$invK)
Mslope=F_basic$coefficients[2]
Tslope=F_basic$coefficients[3]
Ea=Tslope*0.00008617
class_slopes=rbind(class_slopes, c("Actinoperygii", Mslope, Tslope, Ea))
class_slopes=as.data.frame(class_slopes, stringsAsFactors = FALSE)

#removing unneccesary column and renaming columns
class_values=class_slopes[c(-3)]
names(class_values)[names(class_values) == "V1"] = "Class"
names(class_values)[names(class_values) == "log(Class_data$Mg)"] = "exponent"
names(class_values)[names(class_values) == "Class_data$InvK"] = "Ea"

# Classes: Insecta, Gastropoda, Eurotatoria, Entognatha
# Data source (Insecta only): Chown
# Remaining classes' values from theory
I_class="Insecta"
other_class=c("Gastropoda","Eurotatoria", "Entognatha")
I_slope=0.75
I_Ea=-0.62
Other_Ea=-0.63
other_values=cbind(I_class,I_slope,I_Ea)
for (current_class in other_class){
  current_values=cbind(current_class,I_slope,Other_Ea)
  other_values=rbind(other_values, current_values)
}

other_values = as.data.frame(other_values)
names(other_values)[names(other_values) == "I_class"] = "Class"
names(other_values)[names(other_values) == "I_slope"] = "exponent"
names(other_values)[names(other_values) == "I_Ea"] = "Ea"
class_values = rbind(class_values, other_values)
class_values = transform(class_values, exponent = as.numeric(exponent), Ea = as.numeric(Ea))

class_values
```

#### CREATE THREE DEGREE DATASET

Put together replicates that have 3* differences in temperature, removing other replicates

```{r, echo=FALSE}
# Read in compiled experimental dataset, has one temperature and replicate per row
original_data = read.csv("MTEEcto_data.csv", stringsAsFactors = FALSE, fileEncoding = "latin1")

# Turn unique row names into column in original dataset
original_data = data.frame(as.numeric(rownames(original_data)), original_data)
colnames(original_data)[1] = "row_names"

# Create list of experiment names and respective species
replicate_list_original=unique(original_data$studyID)

# Create dataset containing all temperature pairs within each replicate
all_pairs = c()
for(current_replicate in replicate_list_original){
  single_replicate = subset(original_data, original_data$studyID == current_replicate)
  single_replicate_rownames_combos = combn(single_replicate$row_names, 2)
  single_replicate_rownames_combos = t(single_replicate_rownames_combos)
  single_replicate_temps_combos = combn(single_replicate$temp, 2)
  single_replicate_temps_combos = t(single_replicate_temps_combos)
  single_replicate_masses_combos = combn(single_replicate$mass, 2)
  single_replicate_masses_combos = t(single_replicate_masses_combos)
  single_replicate_combos = cbind(single_replicate_rownames_combos, single_replicate_temps_combos, single_replicate_masses_combos)
  single_replicate_combos = data.frame(single_replicate_combos)
  single_replicate_combos = cbind(current_replicate, single_replicate_combos)
  all_pairs = rbind(all_pairs, single_replicate_combos)
}
names(all_pairs) = c("studyID", "initial_temp_row", "final_temp_row", "initial_temp", "final_temp", "initial_mass", "final_mass")

# Subset dataset to contain only temperature pairs that differ by 3*
all_pairs$temp_diff = all_pairs$final_temp - all_pairs$initial_temp
three_degree_pairs = subset(all_pairs, all_pairs$temp_diff == 3)

# Add species to three degree difference pairs
lookup = original_data[match(three_degree_pairs$studyID, original_data$studyID),]
three_degree_pairs$species = lookup$Species
three_degree_pairs$Class = lookup$Class

# List of species and number of pairs per species
species_duplicates = sort(table(three_degree_pairs$species), decreasing=TRUE)

# Add class-specific values to pairs dataset
store_class_values = class_values[match(three_degree_pairs$Class, class_values$Class),]
three_degree_pairs = cbind(three_degree_pairs, store_class_values$exponent, store_class_values$Ea)
colnames(three_degree_pairs)[11] = "exponent"
colnames(three_degree_pairs)[12] = "Ea"

# List of classes and number of pairs per class
class_distribution = table(three_degree_pairs$Class)

head(three_degree_pairs)
```
